---
title: "Homework8"
author: "Akhil Havaldar"
date: "10/30/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readr)
```

## Question 1
```{r}
set.seed(05202001)
delta <- read.csv("Delta delays.csv")
delta <- as.vector(delta$Arrival.Delay)
quants <- quantile(delta)
samp_iqr <- IQR(delta)

B<-10000
boot_samp <- replicate(B, sample(delta, replace=T))
boot_iqr <- apply(boot_samp,2,IQR)

boot_err <- boot_iqr - samp_iqr
boot_err_sort <- sort(boot_err)
p5.0 <- B*0.05
p95.0 <- B*0.95
boot_ci <- samp_iqr - boot_err_sort[c(p95.0,p5.0)]
print(boot_ci)
```
- The bootstrap 90% CI for IQR is (-3.5, 21.0)

## Question 2
### a) 
```{r}
quants <- quantile(delta)
samp_90 <- quantile(delta,c(.9))

B<-10000
boot_samp <- replicate(B, sample(delta, replace=T))
boot_90 <- apply(boot_samp,2,quantile, 0.9)

boot_90 <- boot_90 - samp_90
boot_err_sort <- sort(boot_err)
p025 <- B*0.025
p97.5 <- B*0.975
boot_ci <- samp_iqr - boot_err_sort[c(p97.5,p025)]
print(boot_ci)
```
- We get a value of (-8,22.5) for the CI. There is some evidence that the 90th percentile of arrival delay length is a late arrival. 
### b) 
- No this is not indisputable. For it to be indisputable, the entire confidence interval for the 90th percentile has to be positive. 

## Question 3
```{r}
set.seed(05202001)
amer <- read.csv("American delays.csv")
amer <- as.vector(amer$Arrival.Delay)
amer90 <- quantile(amer, c(.90))
print(amer90)
print(samp_90)

samp_diff <- samp_90 - amer90
print(samp_diff)

rand.test <- function(x){
  rand_comb <- sample( c(delta, amer) )
  bquant1 <- quantile(rand_comb[1:x], c(0.90))
  bquant2 <- quantile(rand_comb[(x+1):(length(delta)+length(amer))], c(0.90))
  bquant1 - bquant2
}

boot_diffs_null <- replicate(B, rand.test(length(delta)))
sum(boot_diffs_null <= samp_diff | boot_diffs_null >= 2*mean(boot_diffs_null) - samp_diff)/B
```
- Fail to Reject null, which means that there is no evidence that the 90th percentile of arrival delay lengths is different between the two airlines.
