---
title: "Homework7"
author: "Akhil Havaldar"
date: "10/24/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
```{r}
set.seed(05202001)
library(ggplot2)

dat <- data.frame(X=c(0,20))
ggplot(dat, aes(x=X)) + stat_function(fun=dchisq, args=list(df=3))
```
- The density curve is heavily right skewed because of its low degrees of freedom.

## Question 2
### 2a) 
```{r}
mu <- 3
sig <- 2.45
alpha <- .05

ztest2 <- function(size) {
  sample <- rchisq(size, df=3)
  samplemean <- mean(sample)
  teststat <- (samplemean - mu) / (sig/sqrt(size))
  pvalue <- 2*pnorm(-(abs(teststat)))
  if(pvalue < alpha){TRUE}
  else{FALSE}
}

K <- 10000
t1two <- function(size){
  samps <- replicate(K, ztest2(size))
  length(which(samps == TRUE)) / 10000
}

q2a <- sapply(c(8,23,52), t1two)
q2a
```

### 2b)
```{r}
ztleft <- function(size) {
  sample <- rchisq(size, df=3)
  samplemean <- mean(sample)
  teststat <- (samplemean - mu) / (sig/sqrt(size))
  pvalue <- pnorm(teststat, lower.tail = TRUE)
  if(pvalue < alpha){TRUE}
  else{FALSE}
}

K <- 10000
t1lft <- function(size){
  samps <- replicate(K, ztleft(size))
  length(which(samps == TRUE)) / 10000
}

q2b <- sapply(c(8,23,52), t1lft)
q2b

```

### 2c) 
```{r}
ztright <- function(size) {
  sample <- rchisq(size, df=3)
  samplemean <- mean(sample)
  teststat <- (samplemean - mu) / (sig/sqrt(size))
  pvalue <- pnorm(teststat, lower.tail = FALSE)
  if(pvalue < alpha){TRUE}
  else{FALSE}
}

K <- 10000
type1right <- function(size){
samps <- replicate(K, ztright(size))
length(which(samps == TRUE)) / 10000
}

q2c <- sapply(c(8,23,52), type1right)
q2c

```

## Question 3
```{r}
out <- data.frame(q2a, q2b, q2c)
rownames(out) <- c("n=8", "n=23", "n=52")
colnames(out) <- c("two-sided","left-sided", "right-sided")
out
```

- As the sample size increases for each test, the type 1 error gets closer and closer to the significance level of 0.05. 

## Question 4
### 4a) 
```{r}
t2 <- function(size){
  sample <- rchisq(size, df=3)
  teststat <- (mean(sample) - mu) / (sd(sample)/sqrt(size))
  pvalue <- 2*pt(-abs(teststat), df=(size-1))
  pvalue < alpha
}

K <- 10000
t1twoT <- function(size){
  samps <- replicate(K, t2(size))
  length(which(samps == TRUE)) / 10000
}

q4a <- sapply(c(8,23,52), t1twoT)
q4a
```

### 4b) 
```{r}
ttleft <- function(size){
  sample <- rchisq(size, df=3)
  teststat <- (mean(sample) - mu) / (sd(sample)/sqrt(size))
  pvalue <- pt(teststat, df=(size-1), lower.tail = TRUE)
  pvalue < alpha
}

K <- 10000
t1lftT <- function(size){
  samps <- replicate(K, ttleft(size))
  length(which(samps == TRUE)) / 10000
}

q4b <- sapply(c(8,23,52), t1lftT)
q4b

```

### 4c) 
```{r}
ttright <- function(size){
  sample <- rchisq(size, df=3)
  teststat <- (mean(sample) - mu) / (sd(sample)/sqrt(size))
  pvalue <- pt(teststat, df=(size-1), lower.tail = FALSE)
  pvalue < alpha
}

K <- 10000
type1rightT <- function(size){
  samps <- replicate(K, ttright(size))
  length(which(samps == TRUE)) / 10000
}

q4c <- sapply(c(8,23,52), type1rightT)
q4c
```

## Question 5
```{r}
out2 <- data.frame(q4a, q4b, q4c)
rownames(out2) <- c("n=8", "n=23", "n=52")
colnames(out2) <- c("two-sided","left-sided", "right-sided")
out2
```

- As the sample size increases for each test, the type 1 error gets closer and closer to the significance level of 0.05. 

## Question 6
### 6a) 
```{r}
zbad <- function(size) {
  sample <- rchisq(size, df=3)
  samplemean <- mean(sample)
  teststat <- (samplemean - mu) / (sd(sample)/sqrt(size))
  pvalue <- 2*pnorm(-(abs(teststat)))
  if(pvalue < alpha){TRUE}
  else{FALSE}
}

K <- 10000
t1twoZbad <- function(size){
  samps <- replicate(K, zbad(size))
  length(which(samps == TRUE)) / 10000
}

q6a <- sapply(c(8,23,52), t1twoZbad)
q6a
```

### 6b) 
```{r}
ztleftbad <- function(size) {
  sample <- rchisq(size, df=3)
  samplemean <- mean(sample)
  teststat <- (samplemean - mu) / (sd(sample)/sqrt(size))
  pvalue <- pnorm(teststat, lower.tail = TRUE)
  if(pvalue < alpha){TRUE}
  else{FALSE}
}

K <- 10000
t1lftZbad <- function(size){
  samps <- replicate(K, ztleftbad(size))
  length(which(samps == TRUE)) / 10000
}

q6b <- sapply(c(8,23,52), t1lftZbad)
q6b
```

### 6c)
```{r}
ztrightbad <- function(size) {
  sample <- rchisq(size, df=3)
  samplemean <- mean(sample)
  teststat <- (samplemean - mu) / (sd(sample)/sqrt(size))
  pvalue <- pnorm(teststat, lower.tail = FALSE)
  if(pvalue < alpha){TRUE}
  else{FALSE}
}

K <- 10000
type1rightZbad <- function(size){
  samps <- replicate(K, ztrightbad(size))
  length(which(samps == TRUE)) / 10000
}

q6c <- sapply(c(8,23,52), type1rightZbad)
q6c
```

### Question 7
```{r}
out3 <- data.frame(q6a, q6b, q6c)
rownames(out3) <- c("n=8", "n=23", "n=52")
colnames(out3) <- c("two-sided","left-sided", "right-sided")
out3
```

- As the sample size increases for each test, the type 1 error gets closer and closer to the significance level of 0.05. 
